Efficient spherical Convolutional Neural Networks with Healpix sampling for cosmological applications

### 2018-04-27 scnn.tex
some more work

\section{Method}

@@

\subsection{Graph creation}
\assign{Nathanaël}

In the classical 2-dimenstional convolution, each pixel is connected with the
same weight to its 4 closest neighbors. We use a weighted undirected graph to
generalize this idea to the ubiquitous HealPix
sampling~\citep{gorski2005healpix}.  Each pixel is represented by a node
(vertex) connected to his $8$ or $7$ closest neighbors.\footnote{For some
pixels, the $8^{th}$ nearest neighbor is not well defined.} Given the set of
nearest neighbors, we define the weight matrix $W$ using the following scheme
\begin{equation}
W[i,j]=\begin{cases}
e^{-\frac{\|x_i-x_j\|_2^2}{\sigma^2}} & \text{if pixels $i$ and $j$ are connected, and}\\
0 & \text{otherwise.}\\
\end{cases}
\end{equation}
Here $x_i$ is a 3-dimensional vector encoding the coordinate of the pixels $i$
on the sphere and $\sigma$ is the mean of $\|x_i-x_j\|_2$ over all connected
pixels $i$ and $j$. This weighting scheme is important as the distances between
pixels varries du to the small sampling irregularities. Some of the
irregularities have been highlighted in the graph displayed in
Figure~\ref{fig:healpix_graph_4}.

These small irregularities are important as they prevent our approach to be affect slightly the degree of a node (or a pixel), defined as $d_i =
\sum_j W[i,j]$. The degree matrix $D$ is the diagonal matrix where $D_{ii}=d_i$.

\begin{figure}[!ht]
\centering
% \includegraphics[width=0.45\textwidth]{figures/healpix_graph_4.pdf}
\includegraphics[width=0.45\textwidth]{figures/half_graph_4.pdf}
\caption{
% Left: Healpix graph for the full sphere using a nside parameter of $4$. 
Top view of a Healpix graph for half the sphere using a nside parameter of $4$.}
\label{fig:healpix_graph_4}
\end{figure}

\subsection{Graph Fourier basis and spherical harmonics}
\assign{Nathanaël}
% TODO: basis or transform?

The graph normalized graph Laplacian defined as $L = I - D^{-1/2} W D^{-1/2}$ is a second order differential operator that can be used to define the graph Fourier basis.

@@

\begin{itemize}
	\item We build a graph using the healpix sampling
	\item Define Fourier transform and show that the harmonics are visually close to the spherical harmonics
	\item Define spherical convolution using the graph Fourier transform and show heat diffusion example
	\item Show the limits of the approach and explain why we cannot have a perfect spherical convolution with this technique
\end{itemize}

@@

With the Fourier basis. Because we don't have an FFT, that costs $O(n^3)$ for the eigendecomposition, plus $O(n^2)$ for the transform (to be done for each forward and backward pass).

@@

\begin{figure}[!ht]
\centering
\includegraphics[width=0.45\textwidth]{figures/index_plotting_order20_nside16.pdf}

\caption{Indexes selected for the section ploting of the filter.}
\label{fig:index_section}
\end{figure}

\subsection{Efficient convolutions}
\assign{Michaël}

\begin{itemize}
	\item We need efficient convolution, hence the Chebysheff trick
	\item Explain the different between a polynomial filter convolution and a traditional patch restricted convolution
	\item Define spherical CNN using graph CNN
\end{itemize}

>>>>

\section{Method}
% \begin{itemize}
% 	\item We build a graph using the healpix sampling
% 	\item Define Fourier transform and show that the harmonics are visually close to the spherical harmonics
% 	\item Define spherical convolution using the graph Fourier transform and show heat diffusion example
% 	\item Show the limits of the approach and explain why we cannot have a perfect spherical convolution with this technique
% \end{itemize}

@@

\subsection{Graph creation}
In the classical 2-dimenstional convolution, each pixel is connected with the
same weight to its 4 closest neighbors. We use a weighted undirected graph to
generalize this idea to the ubiquitous HealPix
sampling~\citep{gorski2005healpix}.  Each pixel is represented by a node
(vertex) connected to his $8$ or $7$ closest neighbors.\footnote{For some
pixels, the $8^{th}$ nearest neighbor is not well defined.} Given the set of
nearest neighbors, we define the weight matrix $W$ using the following scheme
\begin{equation}
W[i,j]=\begin{cases}
e^{-\frac{\|x_i-x_j\|_2^2}{\sigma^2}} & \text{if pixels $i$ and $j$ are connected, and}\\
0 & \text{otherwise.}\\
\end{cases}
\end{equation}
Here $x_i$ is a 3-dimensional vector encoding the coordinate of the pixels $i$
on the sphere and $\sigma$ is the mean of $\|x_i-x_j\|_2$ over all connected
pixels $i$ and $j$. This weighting scheme is important as the distances between
pixels varries du to the small sampling irregularities. Some of the
irregularities have been highlighted in the graph displayed in
Figure~\ref{fig:healpix_graph_4}.

These small irregularities are important and affect slightly the degree of a 
node (or a pixel), defined as $d_i =\sum_j W[i,j]$.

\begin{figure}[!ht]
\centering
% \includegraphics[width=0.45\textwidth]{figures/healpix_graph_4.pdf}
\vspace{-0.5cm}
\includegraphics[width=0.45\textwidth]{figures/half_graph_4.pdf}
\vspace{-0.5cm}
\caption{
% Left: Healpix graph for the full sphere using a nside parameter of $4$. 
Top view of a Healpix graph for half the sphere using a nside parameter of $4$.}
\label{fig:healpix_graph_4}
\end{figure}

\subsection{Graph Fourier basis and spherical harmonics}

Because of the irregular sampling, there is no obvious way to define convolution
directly from the pixel values. Hence we first define a Fourier transform on the
graph.

Following~\cite{shuman2013emerging}, the graph normalized graph Laplacian
defined as $L = I - D^{-1/2} W D^{-1/2}$ is a second order differential operator
that can be used to define the graph Fourier basis. Here $D$ is the diagonal
matrix where $D_{ii}=d_i$. The graph Fourier basis is defined as the harmonic
decomposition of the graph Laplacian $L=U \Lambda U^*$, where $U$ is the
orthonormal matrix of eigenvectors and $\Lambda$, the diagonal matrix of
eigenvalues.

@@

As shown in Figure~\ref{fig:graph_harmonics}, it turns out that using the graph
construction described above, the graph Fourier harmonics ressembles the
classical spherical harmonics giving us an indication that the graph is able to
capture the spherical properties of the HealPix sampling.

@@

Convolution

@@

\begin{figure}[!ht]
\centering
\includegraphics[width=0.45\textwidth]{figures/index_plotting_order20_nside16.pdf}
\caption{Indexes selected for the section ploting of the filter.}
\label{fig:index_section}
\end{figure}

\subsection{Efficient convolutions}
\assign{Michaël}
With the Fourier basis. Because we don't have an FFT, that costs $O(n^3)$ for the eigendecomposition, plus $O(n^2)$ for the transform (to be done for each forward and backward pass).
\begin{itemize}
	\item We need efficient convolution, hence the Chebysheff trick
	\item Explain the different between a polynomial filter convolution and a traditional patch restricted convolution
	\item Define spherical CNN using graph CNN
\end{itemize}
